\bf Title here

### 2020-07-06 paper/paper.Rmd
an attempt to write the three algorithms in math

- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX
- name: Author 2
  affiliation: Department of ZZZ, University of WWW
keywords:
- optimisation
- projection pursuit 
- guided tourr
- visual 
- diagnostics 
- R
abstract: Friedman & Tukey commented on their intial paper on projection pursuit in 1974 that "the  technique  use  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure. In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum. This diagnostic system can be applied to more general to help practitioner to unveil the black-box in randomised iteartive (optimisation) algorithms. An R package, ferrn, has been created to implement this diagnostic system.
bibliography: biblio.bib
preamble: >
  \usepackage{amssymb}
output: rticles::asa_article
editor_options: 
  chunk_output_type: console
---

@@

There are three existing methods for optimisating PPI function and we review them below. @posse1995projection used a stochastic direct search method,  a random search algorithm to sample new bases in the neighbourhood of the current basis defined by the radius of the p-dimensional sphere, $c$. The new basis is taken as the target basis if it has higher index value, or the sampling continues. If no basis is found to have higher index value after a certain number of tries $n$, the radius $c$ is halved. The algorithm stops when the maximum number of iteration is attained or the radius $c$ is less than a pre-determined number.

@@

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $B$, a candidate basis is constructed as $B_j = (1- \alpha)B_i + \alpha B$ where $B_i$ is the current basis. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)$ where $I(.)$ is the index function.

>>>>

- name: Author 1
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of YYY, University of XXX
- name: Author 2
  affiliation: Department of ZZZ, University of WWW
keywords:
- optimisation
- projection pursuit 
- guided tourr
- visual 
- diagnostics 
- R
abstract: Friedman & Tukey commented on their intial paper on projection pursuit in 1974 that "the  technique  use  for  maximising  the  projection index strongly influences both the statistical and the computational aspects of the procedure." While many projection pursuit indices have been proposed in the literature, few concerns the optimisation procedure. In this paper, we developed a system of diagnostics aiming to visually learn how the optimisation procedures find its way towards the optimum. This diagnostic system can be applied to more general to help practitioner to unveil the black-box in randomised iteartive (optimisation) algorithms. An R package, ferrn, has been created to implement this diagnostic system.
bibliography: biblio.bib
preamble: >
  \usepackage{amssymb}
  \usepackage[ruled,vlined, linesnumbered]{algorithm2e}
output: rticles::asa_article
editor_options: 
  chunk_output_type: console
---

@@

There are three existing methods for optimisating PPI function and we review them below.

@posse1995projection used a stochastic direct search method, a random search algorithm to sample a candidate bases $A^{(j +1)}$ in the neighbourhood of the current basis $A^j$ defined by the radius of the p-dimensional sphere, $\alpha$. If the candidate basis has a higher index value than the current basis, it is outputed as the target basis $A^{j + 1}$ along with other metadata and the algorithm stops. If no basis is found to have higher index value after the maximum number of tries $i_{max}$, the algorithm stops with nothing outputted. $h$ is the halfing parameter meaning if $h > 30$, the nieghbourhood parameter $\alpha$ will be reduced by half in the next iteration of the random search. The algorithm for a random search is summarised in Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $h = 0$, $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
    $I^{(j + 1)} = f(A^{(j+1)})$\;
    \eIf{$I^{(j+1)} > I^{j}$}{
      $A^{j+1} = A^{(j+1)}$\;
      }{
      $h = h + 1$\;
      }
    $i = i + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing [@bertsimas1993simulated, @kirkpatrick1983optimization] is a non-derivative procedure based on a non-increasing cooling scheme $T(i)$. Given an initial $T_0$, the temperature at iteration $i$ is defined as $T(i) = \frac{T_0}{log(i + 1)}$. The simulated annealing algorithm works as follows. Given a neighbourhood parameter $\alpha$ and a randomly generated orthonormal basis $A_{rand}$, a candidate basis is constructed as $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$. If the index value of the candidate basis is larger than the one of the current basis, the candidate basis becomes the target basis. If it is smaller, the candidate is accepted with probability $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$ where $I^j$ and $I^{(j+1)}$ are the index value of the current and candidate basis respectively. The algorithm can be summarised as in Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$\;
  \While{$i < i_{max}$}{
    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{(j+1)} = f(A^{(j+1)})$\;
    $T(i) = \frac{T_0}{\log(i + 1)}$\;
      \eIf{$I^{(j+1)} > I^{j}$}{
        $A^{j+1} = A^{(j+1)}$\;
        $I^{j+1} = I^{(j+1)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
        }
      }
    $i = i + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

\newpage

### 2020-07-06 paper/paper.tex
an attempt to write the three algorithms in math

%% load any required packages here
\usepackage{amssymb}

@@

There are three existing methods for optimisating PPI function and we
review them below. \citet{posse1995projection} used a stochastic direct
search method, a random search algorithm to sample new bases in the
neighbourhood of the current basis defined by the radius of the
p-dimensional sphere, \(c\). The new basis is taken as the target basis
if it has higher index value, or the sampling continues. If no basis is
found to have higher index value after a certain number of tries \(n\),
the radius \(c\) is halved. The algorithm stops when the maximum number
of iteration is attained or the radius \(c\) is less than a
pre-determined number.

@@

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(B\), a candidate basis is
constructed as \(B_j = (1- \alpha)B_i + \alpha B\) where \(B_i\) is the
current basis. If the index value of the candidate basis is larger than
the one of the current basis, the candidate basis becomes the target
basis. If it is smaller, the candidate is accepted with probability
\(A = \min \left(\exp(-\frac{I(B_j) - I(B_i)}{T(i)}), 1 \right)\) where
\(I(.)\) is the index function.

>>>>

%% load any required packages here
\usepackage{amssymb} \usepackage[ruled,vlined, linesnumbered]{algorithm2e}

@@

There are three existing methods for optimisating PPI function and we
review them below.

\citet{posse1995projection} used a stochastic direct search method, a
random search algorithm to sample a candidate bases \(A^{(j +1)}\) in
the neighbourhood of the current basis \(A^j\) defined by the radius of
the p-dimensional sphere, \(\alpha\). If the candidate basis has a
higher index value than the current basis, it is outputed as the target
basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
If no basis is found to have higher index value after the maximum number
of tries \(i_{max}\), the algorithm stops with nothing outputted. \(h\)
is the halfing parameter meaning if \(h > 30\), the nieghbourhood
parameter \(\alpha\) will be reduced by half in the next iteration of
the random search. The algorithm for a random search is summarised in
Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $h = 0$, $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
    $I^{(j + 1)} = f(A^{(j+1)})$\;
    \eIf{$I^{(j+1)} > I^{j}$}{
      $A^{j+1} = A^{(j+1)}$\;
      }{
      $h = h + 1$\;
      }
    $i = i + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(A_{rand}\), a candidate
basis is constructed as
\(A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
of the candidate basis is larger than the one of the current basis, the
candidate basis becomes the target basis. If it is smaller, the
candidate is accepted with probability
\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}\)
where \(I^j\) and \(I^{(j+1)}\) are the index value of the current and
candidate basis respectively. The algorithm can be summarised as in
Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$\;
  \While{$i < i_{max}$}{
    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{(j+1)} = f(A^{(j+1)})$\;
    $T(i) = \frac{T_0}{\log(i + 1)}$\;
      \eIf{$I^{(j+1)} > I^{j}$}{
        $A^{j+1} = A^{(j+1)}$\;
        $I^{j+1} = I^{(j+1)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
        }
      }
    $i = i + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

\newpage
